Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data

### 2022-11-11 paper/manuscript.Rmd
AB text edits

In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available NHANES data. Using this application, we generated crude and age-adjusted BP and hypertension statistics. Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women. Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.

@@

NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension. SBP and DBP were measured following a standardized protocol by trained and certified physicians.  NHANES data are publicly available to download, but working with these data requires understanding variable definitions and advanced programming and statistical knowledge. Challenges that users encounter when analyzing NHANES data include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data. The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.

@@

We were able to replicate several prior manuscripts using the application.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%. The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes. The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist. We included spironolactone as a single drug class, an aldosterone antagonist. Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications. Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.

>>>>

In the current manuscript, we present a web application with a user-friendly point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using publicly available NHANES data. Using this application, we generated crude and age-adjusted BP and hypertension statistics. Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women. Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults to inform their research and policy decisions.

@@

NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension. SBP and DBP were measured following a standardized protocol by trained and certified physicians.  NHANES data are publicly available to download, but working with these data requires understanding variable definitions and advanced programming and statistical knowledge. Challenges of analyzing NHANES data may also include downloading and merging multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data. The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.

@@

Using the application, we replicated results from several prior manuscripts.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%. The difference in the prevalence estimates can be attributed to the approach used to categorize antiyhpertensive medications into their medication classes. The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist. We included spironolactone as a single drug class, an aldosterone antagonist. Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications. Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.
